---
title: "SupplementaryInfo_GibsonHay_LaS"
author: "Andy Gibson, Jen Hay"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# The *bad* sounds *bed*-er in a song: New Zealand listeners expect American singing accents

## Supplementary Materials
Andy Gibson and Jen Hay

## Contents
- This R Markdown file contains the code used for running the three models presented in the article, as well as for generating the plots.

1. Read in Experiment 1 (PCT) data
1. Run Experiment 1 model
1. Plot Experiment 1 model predictions
1. Read in Experiment 2 (LDT) data
1. Run Experiment 2 accuracy model
1. Run Experiment 2 reaction time model
1. Plot Experiment 2 reaction time model predictions

## Load libraries and read in Experiment 1 data

```{r load libraries}
### load libraries ====
library(tidyverse)
library(lme4)
library(effects)
library(emmeans)

```



```{r read in PCT data}
read.csv()


```

## Run Experiment 1 model

## Plot Experiment 1 model predictions


## Read in Experiment 2 (LDT) data

## Run Experiment 2 accuracy model

## Run Experiment 2 reaction time model

## Plot Experiment 2 reaction time model predictions



